Modeling the Java Card API (or Interface?) in B

### 2012-03-23 trunk/papers/B2012-work/javacardAPIpaper.tex
alguns pequenos ajustes e tarefas para Simone.

\begin{document}
\begin{frontmatter}
  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{Simone - Bruno - David \thanksref{coemail}}
  \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
        anamaria@dimap.ufrn.br}}} \thanks[coemail]{Email:
    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
        couserid@codept.coinst.coedu}}}

% AMM: Precisa ser arrumado: The ENTCS macro package provides two alternatives to listing authors
% names and addresses.  These are described in detail in the file
% \texttt{instraut.dvi}. Basically, listing each author and his or her
% address in turn, is the simplest method.  But, if there are several
% authors and two or more share the same address (but not all authors
% are at this address), then the method of listing authors first, and
% then the addresses, and of referencing addresses to authors should be
% used.

@@

In the context of the BSmart project, we are interested in the use of
B to develop Java Card~\cite{chen:2000} applications.  Java Card is a
restricted and optimized version of Java which allows memory and
processor constrained devices, such as smart cards, to store and run
small applications.  Java Card applications have characteristics
(small size and need for security) which make then good candidates for
formal development with the B method. They also present some
regularity of structure that induced our group into the definition of
a method and a tool which specialize B for development of such
applications, providing greater support for the developer than pure B
tools \cite{atelierB, proB}. They are the BSmart method and tool
presented in \cite{mainpapers}.

Previous works have also dealt with B for Java Card, \cite{BOM,
  JBTools e cia}, but none has, at our knowledge, specified Java Card
in B so that the formal development includes Java Card aspects. In the
first versions of BSmart \cite{}, these aspects were included by the code
ganerator, and in JBTools \cite{}, the generated code needed to be edited by
the developer to include these aspects.

These works lacked then the ability to formally verify the Java Card
dependent aspects of the application. To be able to change this, a B
specification of these aspects is needed, and this contribution is
what we present in this paper: the B specification of the Java Card
API, with...  This specification can be used with the BSmart tool,
independently, in any B tool, or even as a complement to the official
Java Card documentation \cite{}.

@@

>>>>

\begin{document}
\begin{frontmatter}
%  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{S. de Oliveira Santos \thanksref{simone}}
%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{B. Emerson Gurgel Gomes \thanksref{bruno}}
%    \address{IFRN, Currais Novos, RN, Brazil} \author{D. Boris Paul D\'{e}harbe \thanksref{david}}
%      \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
%    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
%        anamaria@dimap.ufrn.br}}} 
%      \thanks[simone]{Email:
%    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
%        simone82@ppgsc.ufrn.br}}}
%      \thanks[bruno]{Email:
%    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
%        bruno.gurgel@ifrn.edu.br}}}
%       \thanks[david]{Email:
%    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
%        david@dimap.ufrn.br}}}

  \title{Modeling the Java Card API (or Interface?) in B} \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
  \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}}
  \author[ufrn]{D.  D\'{e}harbe \thanksref{david}}
  \address[ufrn]{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais? REUNI?}  \address[ifrn]{IFRN, Currais Novos, RN, Brazil} 
 \thanks[anamaria]{Email:
    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
        anamaria@dimap.ufrn.br}}} 
      \thanks[simone]{Email:
    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
        simone82@ppgsc.ufrn.br}}}
      \thanks[bruno]{Email:
    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
        bruno.gurgel@ifrn.edu.br}}}
       \thanks[david]{Email:
    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
        david@dimap.ufrn.br}}}

@@

In the context of the BSmart project, we are interested in the use of
B to develop Java Card~\cite{chen:2000} applications.  Java Card is a
restricted and optimized version of Java which allows memory and
processor constrained devices, such as smart cards, to store and run
small applications.  Java Card applications have characteristics
(small size and need for security) which make then good candidates for
formal development with the B method. They also present some
regularity of structure that induced our group into the definition of
a method and a tool which specialize B for development of such
applications, providing greater support for the developer than pure B
tools \cite{atelierB,proB}. They are the BSmart method and tool,
presented in \cite{BSmart08,Gomes10}.

Previous works have also dealt with B for  Smart Cards, 
\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at our knowledge, specified Java Card
in B so that the formal development includes Java Card aspects. In the
first versions of BSmart \cite{Gomes10}, these aspects were included by the code
ganerator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
the developer to include these aspects.

These works lacked then the ability to formally verify the Java Card dependent aspects of the application. To be able to change this, a B
specification of these aspects is needed, and this contribution is what we present in this paper: the B specification of the Java Card
API, with its primitive types and all classes of the 14 packages which define the API. This specification can be used with the BSmart tool,
independently, in any B tool, or even as a complement to the official Java Card documentation \cite{oracle}.

SIMONE -> verifique que está certo o que falei acima sobre a especificação.

@@

Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
Suppose, for instance, that we had the B abstract machine of fig \ref{maquina}, which models ...  Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.  As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine. However,  Java Card code corresponding to it would look like the one presented in fig. \ref{impl java card}. In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code. So, we need to be able to refine the original B machine into something such as the B refinement of fig. \ref{B ref usando a API}. Then, code generation would be straightforward, as it should be. This is then the reason why it needed to be the first Java Card component to be modeled in B. (AMM: maybe also cite related work ? WHich are the formalization works that did not focus on the API?)

SIMONE -> por favor proponha um bem pequeno exemplo para colocarmos aqui. A idéia é que mostre que ser B0 não basta para gerar o código Java Card.

The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in table \ref{incluir a tabela da dissert da simone}. Their interfaces have all been modeled in B, and the resulting models are available at LINK. Differently from  JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.  We present in the following the main characteristics and restrictions of our models:

SIMONE-> coloque a tabela, por favor. o LINK pode ser substituído pela página do google code.

\begin{itemize}

\item Only the abstract models corresponding to each class or interface have been developed. Two reasons led to this choice: (1) code is not to be generated from the specification of the API, as we are only creating an abstract layer to existing implementations; and (2)  in B, only the abstract machines are seen by modules that build upon others, through the SEES, USES, IMPORTS, INCLUDES and EXTENDS clauses.

\item Conditions on method inputs which lead to exception raising in the Java Card platform specification are coded into pre-conditions in B.
(em que casos isso não é feito - NULL? outros?)

\item Method outputs are modeled in B as taking any value of the correct type (e em que casos temos um pouco mais).

\end{itemize}
